### 2016-10-12 ID_supp.tex
a

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfigure}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{A Noise Estimation Free Framework for Robust Real Image Denoising}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

 %%%%%%%%% BODY TEXT
\begin{figure*}
\centering
\subfigure{
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Noisy_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (a) Noisy Image \\ (35.05dB/0.9486) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Mean_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (b) Mean Image}
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_CBM3D_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (c) CBM3D \\ (35.06dB/0.9491)}
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_CSF_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (d) CSF \\ (33.17dB/0.9545)}
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_MLP_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (e) MLP \\ (33.04dB/0.9492)}
\end{minipage}
}
\subfigure{
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_NC_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (f) Noise Clinic \\ (36.62dB/0.9743) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_NI_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (g) Neat Image \\ (35.87dB/0.9656) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Offline_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (h) Offline \\ (35.14dB/0.9715)}
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Online_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (i) Online \\ (36.29dB/0.9763)}
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Guided_CC_Noisy_Canon_EOS_5D_Mark3_ISO_3200_C1_52.png}}
{\footnotesize (j) Guided \\ (\textbf{36.95}dB/\textbf{0.9785})}
\end{minipage}
}
\caption{Denoised images of the old image "$Canon EOS 5D Mark3 ISO 3200 C1$" by different methods. The images are better to be zoomed in on screen.}
\label{fig1}
\end{figure*}

\begin{figure*}
\centering
\subfigure{
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Noisy_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (a) Noisy Image \\ (35.92dB/0.8998) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Mean_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (b) Mean Image}
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_CBM3D_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (c) CBM3D \\ (35.94dB/0.9008) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_CSF_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (d) CSF \\ (38.24dB/0.9670) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_MLP_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (e) MLP \\ (36.38dB/0.9539)}
\end{minipage}
}
\subfigure{
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_NC_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (f) Noise Clinic \\ (39.18dB/0.9668) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_NI_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (g) Neat Image \\ (38.58dB/0.9534) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Offline_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (h) Offline \\ (39.78dB/0.9770)}
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Online_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (i) Online \\ (39.33dB/0.9747) }
\end{minipage}
\begin{minipage}[t]{0.2\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/br_Guided_CC_Noisy_Nikon_D800_ISO_1600_B3_82.png}}
{\footnotesize (j) Guided \\ (\textbf{39.91}dB/\textbf{0.9761})}
\end{minipage}
}
\caption{Denoised images of the old image "$Nikon D800 ISO 1600 B3$" by different methods. The images are better to be zoomed in on screen.}
\label{fig1}
\end{figure*}
{\small
\bibliographystyle{unsrt}
\bibliography{egbib}
}

\end{document}